---
title: "Class 12 Lab Session"
author: "Irene Hsieh"
format: pdf
---

# Section 1. proportion og G/G in a population

Downloaded a CSv file from Ensemble, Here we read this CSV file

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)
```

```{r}
table(mxl$Genotype..forward.strand.)
```


```{r}
table(mxl$Genotype..forward.strand.)/ nrow(mxl)*100
```

Now let's look at a different population. I picked GBR 

```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(gbr)
```

Find proportion of GG
```{r}
round(table(gbr$Genotype..forward.strand.)/ nrow(gbr)*100,2 )
```

This variant that is associated with childhood asthma is more frequent in the GBR population than the MKL population

Lets now dig into this further. 



## Section 4
> How many samples do we have?

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6")
head(expr)
```

```{r}
nrow(expr)
```

```{r}
table(expr$geno)
```

```{r}
library(ggplot2)
```

Lets make a boxplot 
```{r}
ggplot(expr) +aes(geno,exp, fill=geno)+
  geom_boxplot(notch=TRUE)
```


